### 2015-06-29 doc/thesis/system.tex
Starting to update typing rules after backport
Updating strict versus default
Minor fixes to chapter 4

We present the subtyping rules as a deduction system for the
subtyping relation $\senv \vdash X \subtype Y$, where $X$ and $Y$
are two types of the same kind.
The variable $\senv$ is a set of pairs of recursion variables.
We need this set to record the hypotheses that we assume when checking
recursive types.

@@

The rule \textsc{S-VARARG1} handles the case where both tuple types end
with variadic types:
\[
\begin{array}{c}
\mylabel{S-VARARG2}\\
\dfrac{\senv \vdash F_{1} \cup \Nil \subtype F_{2} \cup \Nil}
      {\senv \vdash F_{1}{*} \subtype F_{2}{*}}
\end{array}
\]

@@

The other rules handle the cases where only one tuple type ends with a variadic type:
\[
\begin{array}{c}
\begin{array}{c}
\mylabel{S-VARARG2}\\
\dfrac{\senv \vdash F_{1} \cup \Nil \subtype F_{2} \;\;\;
       \senv \vdash F_{1}{*} \subtype P_{2}}
      {\senv \vdash F_{1}{*} \subtype F_{2} \times P_{2}}
\end{array}
\;
\begin{array}{c}
\mylabel{S-VARARG3}\\
\dfrac{\senv \vdash F_{1} \subtype F_{2} \cup \Nil \;\;\;
       \senv \vdash P_{1} \subtype F_{2}{*}}
      {\senv \vdash F_{1} \times P_{1} \subtype F_{2}{*}}
\end{array}
\end{array}
\]

These rules state the conditions when variadic tuples of different
length are compatible.
In the next section we will show that we use the subtyping rules for variadic types,
along with the types $\Value$ and $\Nil$, to make our type system reflect
the semantics of Lua on discarding extra parameters and
replacing missing parameters.

@@

The rule \textsc{S-TABLE2} allows width subtyping and covariant keys.
It allows covariant keys because we also want to use \emph{unique}
table types as a way to join table fields that inhabit \emph{closed} table types.
For instance, we want to use the table constructor to initialize
a variable with a table type that describes a hash.

@@

Even though the dynamic type $\Any$ does not interact with subtyping,
it does interact with consistent-subtyping.
We present the consistent-subtyping rules as a deduction system for
the consistent-subtyping relation $\senv \vdash X \lesssim Y$,
where $X$ and $Y$ are two types of the same kind.
As in the subtyping relation, $\senv$ is also a set of pairs of
recursion variables.
We define the consistent-subtyping rules for the dynamic type $\Any$
as follows:
\[
\begin{array}{c}
\begin{array}{c}
\mylabel{C-ANY1}\\
\senv \vdash F \lesssim \Any
\end{array}
\;
\begin{array}{c}
\mylabel{C-ANY2}\\
\senv \vdash \Any \lesssim F
\end{array}
\end{array}
\]

@@

\subsection{Assignment and function application}
\label{sec:assignment}

@@

\subsection{Tables and refinement}
\label{sec:refinement}

@@

The rule \textsc{T-REFINE} uses the auxiliary function \emph{vt}
to adjust the type of the value according to the type of the key.
More precisely, \emph{vt} includes the type $\Nil$ in the type of
the value when the type of the key is not a literal type.
It is also this function that uses \emph{fix} to prevent
\emph{unique} and \emph{open} table types to appear in the type
of the values.

@@

\subsection{Projections}
\label{sec:fap}

>>>>

We present the subtyping rules as a deduction system for the
subtyping relation $\senv \vdash T_{1} \subtype T_{2}$,
where $T_{1}$ and $T_{2}$ are two types of the same kind.
The variable $\senv$ is a set of pairs of recursion variables.
We need this set to record the hypotheses that we assume when checking
recursive types.

@@

The rule \textsc{S-VARARG1} handles subtyping between two
variadic types:
\[
\begin{array}{c}
\mylabel{S-VARARG1}\\
\dfrac{\senv \vdash F_{1} \cup \Nil \subtype F_{2} \cup \Nil}
      {\senv \vdash F_{1}{*} \subtype F_{2}{*}}
\end{array}
\]

@@

The other rules handle subtyping between a varidic type and a pair type:
\[
\begin{array}{c}
\begin{array}{c}
\mylabel{S-VARARG2}\\
\dfrac{\senv \vdash F_{1} \cup \Nil \subtype F_{2} \;\;\;
       \senv \vdash F_{1}{*} \subtype P_{2}}
      {\senv \vdash F_{1}{*} \subtype F_{2} \times P_{2}}
\end{array}
\;
\begin{array}{c}
\mylabel{S-VARARG3}\\
\dfrac{\senv \vdash F_{1} \subtype F_{2} \cup \Nil \;\;\;
       \senv \vdash P_{1} \subtype F_{2}{*}}
      {\senv \vdash F_{1} \times P_{1} \subtype F_{2}{*}}
\end{array}
\end{array}
\]

These rules state the conditions when tuple types of different
length are compatible.
In the next section we will show that we use the subtyping rules for variadic types,
along with the types $\Value$ and $\Nil$, to make our type system reflect
the semantics of Lua on discarding extra parameters and
replacing missing parameters.

@@

The rule \textsc{S-TABLE2} allows width subtyping and covariant keys.
It allows covariant keys because we also want to use \emph{unique}
table types as a way to join table fields that inhabit \emph{closed} table types.
For instance, we want to use the table constructor to initialize
a variable with a table type that describes a hash.
More precisely, this rule states that it is safe to recast the table type
$\{``x":\Integer, ``y":\Integer, ``z":\Integer\}_{unique}$ to
$\{\String:\Integer \cup \Nil\}_{closed}$, as long as the
new type becomes inacessible with the original type.

@@

Even though the dynamic type $\Any$ does not interact with subtyping,
it does interact with consistent-subtyping.
We present the consistent-subtyping rules as a deduction system for
the consistent-subtyping relation $\senv \vdash T_{1} \lesssim T_{2}$,
where $T_{1}$ and $T_{2}$ are two types of the same kind.
As in the subtyping relation, $\senv$ is also a set of pairs of
recursion variables.
We define the consistent-subtyping rules for the dynamic type $\Any$
as follows:
\[
\begin{array}{c}
\begin{array}{c}
\mylabel{C-ANY1}\\
\senv \vdash F \lesssim \Any
\end{array}
\;
\begin{array}{c}
\mylabel{C-ANY2}\\
\senv \vdash \Any \lesssim F
\end{array}
\end{array}
\]

@@

\subsection*{Assignment and function application}
\label{sec:assignment}

@@

\subsection*{Tables and refinement}
\label{sec:refinement}

@@

The rule \textsc{T-REFINE1} uses the auxiliary function \emph{vt}
to adjust the type of the value according to the type of the key.
More precisely, \emph{vt} includes the type $\Nil$ in the type of
the value when the type of the key is not a literal type.
It is also this function that uses \emph{fix} to prevent
\emph{unique} and \emph{open} table types to appear in the type
of the values.

@@

\subsection*{Projections}
\label{sec:fap}
